---
title: "Intro to R"
author: "Joel Herndon & John Little, Data and Visualization Services"
date: "10/04/2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
usePackage <- function(p) {
    if (!is.element(p, installed.packages()[,1]))
        install.packages(p, dep = TRUE)
    require(p, character.only = TRUE, warn.conflicts = FALSE, quietly = TRUE)
}
usePackage("dplyr")
usePackage("readr")
usePackage("ggplot2")
```

## Introduction

(Powerpoint in "slides" folder: *fallR_oct_4_2016.pptx*)

1. Welcome
2. What is r
3. What is [RStudio](https://www.rstudio.com/)
4. What is a Project
5. What is [Docker](https://vm-manage.oit.duke.edu/containers/rstudio)


## Loading and Reviewing Data

For the workshop, we'll use a derivitive of the [mtcars](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html) dataset. `read_csv` is a convenience function that allows R to read data files with comma separated values.  It is part of a library of functions that you can add to R using the `readr` package that we loaded earlier in the class.

### Reading CSV files
```{r get data}
cars <- read_csv("data/cars.csv")
```

### Visually inspecting your data with Rstudio's "Viewer""

R and Rstudio have several different ways that you can view the data loaded into the system. One of the more "excel like" views is the "Viewer" which is built into Rstudio.

```{r viewer}
View(cars)
```

#### Notes on the viewer
* The view is mostly for quick confirmation - it doesn't leave a record in the audit trail.
* Environment Pane has information about data objects
* View output has sorting 

   
### Evaluate the structure of your data
Looking at a data frame's structure is another quick way to get a better understanding of the data you have loaded into a data frame.

```{r structure}
str(cars)
```

#### Similar Commands
R has several other commands that provide information about data objects:

* `tbl_df()`    # This creates a "tibble" or modern data frame - it's easier to read in the console
* `dim()`       # dimensions - provides the number of rows and columns
* `class()`     # data type 

***
  
***

### Exercise Part One: Loading Data
All excercises are inside the "exercises" folder.
Complete Student_Assignments.Rmd

***

## Data management and cleaning with `dplyr`

`dplyr` is a widely adopted r package and makes it easy to use English verbs to manage and clean your data.

Five verbs make it easy to transform and analyze your data.

### Arrange

1. Use **Arrange** to sort rows
```{r arrange}
arrange(cars, cyl, desc(mpg), desc(hp), wt)
```

### Filter

2. Use **Filter** to select rows
```{r filter}
sixcyls <- filter(cars, cyl == 6)
```

### Mutate

3. Use **Mutate** to create new variables (add columns).
```{r mutate}
sixcyls <- mutate(sixcyls, dispWt = disp / wt)
```


### Select

4. Use **Select** to choose columns (variables)
```{r select}
sixcyls <- select(sixcyls, mpg:wt)
```

### Count

5. Use **Count** to summarize observations (or rows)
```{r count}
count(sixcyls, hp)
```

***

### Exercise Part Two: Data Management
Now we'll use some of the dplyr verbs on a dataset about Duke's newest dining facility, the West Union.  Data are located in the data folder for this class with the name **westUnion.csv**.
Once again All excercises are inside the "exercises" folder.
Complete `Student_Assignments.Rmd` Part Two


***

### Combining Commands (magrittr)
One of the great strengths of R is the ability to chain together commands to achieve complex tasks.  For today's workshop, we won't dwell on this, but the following example demonstrates how you can chain or "pipe" `dplyr` commands together to produce complex data workflows.  If you'd like to read more about how this works, check out (magrittr)[https://cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html].
```{r alltogether}

filter(cars, cyl >= 4, cyl <= 6) %>% 
  select(makeModel, mpg:wt) %>%
  mutate(dispWt = disp / wt) %>% 
  arrange(desc(mpg), wt)
```

## Basic Vis with base R (if time)

When in the data exploratory phase, you can quickly generate very basic visualizations with minimal effort.  Your visualizations then appear in the Viewer Pane.  While you may not want to share these visualizations, the simplicity of generating quick histograms can help you understand your data as you begin thinking about your analysis.

Type these commands into your console.

To generate

* Histogram:  `hist(cars$mpg)`
* Scatterplot: `plot(cars$wt, cars$mpg)`
* Barplot:  `barplot(table(cars$cyl))`


## Visualizations via `ggplot2`
Hadley Wickham's `ggplot2` provides an alternate to base R graphics for producing data visualizations.  While the commands may initially seem complex, they follow a standard syntax that often makes it easier to produce compelling visuals using only a few lines of code.  Find out more at [ggplot2](http://ggplot2.org/).  

### ggplot2 Scatter Plot
```{r ggplot}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point(shape = 1) + 
  geom_smooth(method = lm)
```

[more info on ggplot scatterplots](http://www.cookbook-r.com/Graphs/Scatterplots_(ggplot2)/)


### ggplot2 Bar Graph
```{r ggbarplot}
ggplot(data=cars, aes(x = cyl)) +
    geom_bar(stat="bin")
```

```{r ggbarplot2}
mutate(cars, cyl = factor(cyl, labels = c("four","six","eight"))) %>% 
  ggplot(aes(x = cyl)) +
    geom_bar(stat="count")
```


[More Information on ggplot bar and line graphs](http://www.cookbook-r.com/Graphs/Bar_and_line_graphs_(ggplot2)/)

### ggplot2 Histogram
```{r gghist}
ggplot(cars, aes(x = mpg)) +
    geom_histogram(binwidth=1, colour="black", fill="white")
```

[More info on ggplot histograms](http://www.cookbook-r.com/Graphs/Plotting_distributions_(ggplot2))

### ggplot2 Line graph
```{r linegraph}
ggplot(data=cars, aes(x = mpg, y = wt, group = cyl, colour = cyl)) +
    geom_line() #+
    #geom_point()
```

[More info on line graphs](http://www.cookbook-r.com/Graphs/Bar_and_line_graphs_(ggplot2)/)

