---
title: "Intro to R by Data & Visualization Services"
author: "Joel Herndon & John Little"
date: "10/04/2016"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    theme: readable
    highlight: pygments
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
usePackage <- function(p) {
    if (!is.element(p, installed.packages()[,1]))
        install.packages(p, dep = TRUE)
    require(p, character.only = TRUE)
}
```


# Intro to R


## Introduction (Powerpoint in "slides" folder: fallR_oct_4_2016.pptx)

#### Joel - PPT

1. Welcome
2. What is r
3. What is [RStudio](https://www.rstudio.com/)
4. What is a Project
5. What is [Docker](https://vm-manage.oit.duke.edu/containers/rstudio)


The following code block uses a nonstandard way to load packages - if you have trouble, substitute `require` for `usePackage`.

```{r load-packages}
usePackage("dplyr")
usePackage("readr")
usePackage("ggplot2")
```

## Data Management

We'll use a derivitive of the [mtcars](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html) dataset.  Use the `?` to get the codebook. `read_csv` is a convenience function that allows R to read data files with comma separated values.  It is part of a library of functions that you can add to R using the `readr` package that we loaded earlier in the class.

### Load Data
```{r get data}
cars <- read_csv("data/cars.csv")
```

### View your data in a grid

R and Rstudio have several different ways that you can view the data loaded into the system. One of the more "excel like" views is the "Viewer" which is built into Rstudio.

```{r viewer}
View(cars)
```

#### Notes on the viewer
* The view is mostly for quick confirmation - it doesn't leave a record in the audit trail.
* Environment Pane has information about data objects
* View output has sorting 


### Structure of Data Object
Looking at a data frame's structure is another quick way to get a better understanding of the data you have loaded into a data frame.

```{r structure}
str(cars)
```

#### Similar Commands
R has several other commands that provide information about data objects:

* `tbl_df()`    # This creates a "tibble" or modern data frame - it's easier to read
* `dim()`       # dimensions - provides the number of rows and columns
* `class()`     # data type 

## EXERCISE, Part One: Loading Data
All excercises are inside the "exercises" folder.

#### Joel

1. Student_Assignments.Rmd::PART ONE
**Loading Data**

## Data management and cleaning with `dplyr`

`dplyr` is a widely adopted r package and makes it easy to use English verbs to manage and clean your data.

Five English verbs make it easy to transform and analyze your data.

### Arrange

**Arrange** to sort rows
```{r arrange}
arrange(cars, cyl, desc(mpg), desc(hp), wt)
```

### Filter

**Filter** to select rows
```{r filter}
sixcyls <- filter(cars, cyl == 6)
```

### Mutate

**Mutate** to create new variables (add columns).
```{r mutate}
sixcyls <- mutate(sixcyls, dispWt = disp / wt)
```


### Select
**Select** to choose columns (variables)
```{r select}
sixcyls <- select(sixcyls, mpg:wt)
```

### Count
**Count** to summarize observations (or rows)
```{r count}
count(sixcyls, hp)
```


## EXERCISE, Part 2

2. Student_Assignments.Rmd::PART TWO
**Data Management**

1. `arrange()`
2. `filter()`
3. `mutate()`
4. `select()`
5. `count()`

#### Joel

### Put it all together
Using **magrittr** to combine commands
```{r alltogether}

filter(cars, cyl >= 4, cyl <= 6) %>% 
  select(makeModel, mpg:wt) %>%
  mutate(dispWt = disp / wt) %>% 
  arrange(desc(mpg), wt)
```

## Basic Vis with base R (if time)

When in the data exploratory phase, you can quickly generate very basic visualizations with minimal effort.  Your visualizations then appear in the Viewer Pane.  While you may not want to share these visualizations, the simplicity of generating quick histograms can help you understand your data as you begin thinking about your analysis.

Type these commands into your console.

To generate

* Histogram:  `hist(cars$mpg)`
* Scatterplot: `plot(cars$wt, cars$mpg)`
* Barplot:  `barplot(table(cars$cyl))`


## Visualizations via `ggplot2`
Hadley Wickham's `ggplot2` provides an alternate to base R graphics for producing data visualizations.  While the commands may initially seem complex, they follow a standard syntax that often makes it easier to produce compelling visuals using only a few lines of code.  Find out more at [ggplot2](http://ggplot2.org/).  

### ggplot2 Scatter Plot
```{r ggplot}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point(shape = 1) + 
  geom_smooth(method = lm)
```

[more info on ggplot scatterplots](http://www.cookbook-r.com/Graphs/Scatterplots_(ggplot2)/)


### ggplot2 Bar Graph
```{r ggbarplot}
ggplot(data=cars, aes(x = cyl)) +
    geom_bar(stat="bin")
```

```{r ggbarplot2}
mutate(cars, cyl = factor(cyl, labels = c("four","six","eight"))) %>% 
  ggplot(aes(x = cyl)) +
    geom_bar(stat="count")
```


[More Information on ggplot bar and line graphs](http://www.cookbook-r.com/Graphs/Bar_and_line_graphs_(ggplot2)/)

### ggplot2 Histogram
```{r gghist}
ggplot(cars, aes(x = mpg)) +
    geom_histogram(binwidth=1, colour="black", fill="white")
```

[More info on ggplot histograms](http://www.cookbook-r.com/Graphs/Plotting_distributions_(ggplot2))

### ggplot2 Line graph
```{r linegraph}
ggplot(data=cars, aes(x = mpg, y = wt, group = cyl, colour = cyl)) +
    geom_line() #+
    #geom_point()
```

[More info on line graphs](http://www.cookbook-r.com/Graphs/Bar_and_line_graphs_(ggplot2)/)

