---
title: "Stuf-Class"
author: "john"
date: "9/1/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Intro to R


## Housekeeping

[Joel]

* whatever you want to say about R/RStudio

e.g.

1. welcom
2. what is r
3. what is RStudio
4. What is a Project

```{r load-packages}
require(rmarkdown)
require(dplyr)
require(ggplot2)
```



## Data Management

### Load Data
```{r get data}
?mtcars                # To get help, use precede command with '?' 
data("mtcars")         # load specified data set
```

##### Mention ??

* Import Dataset 

### View Data in a grid
```{r view}
View(mtcars)
```
##### Mention  ??

* Environment Pane
* Sorting
* A convenience but not "reproducible"

### Structure of Data Object

```{r structure}
str(mtcars)
```

#### Mention ??
* tbl_df()
* dim()
* class()


### Missing Data

* is.na()
* anyNA()
* na.omit()

## Data Reshaping 

### Arrange

```{r arrange}
arrange(mtcars, cyl, desc(mpg), desc(hp), wt)
```

### Filter

```{r filter}
sixcyls <- filter(mtcars, cyl == 6)
```

> Exercises:  Show sixcyls grid in the source pane.  how many observations?  how many variables?

### Mutate

Add columns
```{r mutatate}
sixcyls <- mutate(sixcyls, dispWt = disp / wt)
```


### Select

### Count

### Put it all together


## Basic Vis


## Better Vis via ggplot2

